---
title: "Prediction using liear regression model"
author: "Koki Ando"
date: "3/19/2018"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: cosmo
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Preparation
### Required packages

```{r}
library(data.table)
library(broom)
library(corrplot)
library(tidyverse)
library(PerformanceAnalytics)
library(GGally)
library(plotly)
```

### Data Preparation
```{r}
salary.table <- 
  fread("https://s3-ap-southeast-2.amazonaws.com/koki25ando/NBA_season1718_salary.csv", 
        data.table = FALSE)
ss <- fread("https://s3-ap-southeast-2.amazonaws.com/playerinfomation/Seasons_Stats.csv",
            data.table = FALSE)
```

#### Data Cleaning
```{r}
stats17 <- 
  ss %>% filter(Year >= 2017)

stats17 <- 
  stats17 %>% 
  select(Year:G, MP, PER, FG:PTS) %>% 
  distinct(Player, .keep_all = TRUE)

stats17 <- stats17 %>% 
  mutate(MPG = MP/G, PPG = PTS/G, APG = AST/G, 
         RPG = TRB/G, TOPG = TOV/G, BPG = BLK/G, 
        SPG = STL/G) 
```
#### Merging Data
```{r}
stats_salary <- merge(stats17, salary.table, by.x = "Player", by.y = "Player")
names(stats_salary)[40] <- "salary17_18"
stats_salary <- stats_salary[-39]
```

## Correlation check
### Correlation check 1
```{r}
corrplot(cor(stats_salary %>% 
               select(salary17_18, MPG:SPG, 
                      Age, PER, contains("%")), 
             use = "complete.obs"), 
         method = "circle",type = "upper")
```

### Correlation check 2
```{r}
stats_salary_cor <- 
  stats_salary %>% 
  select(salary17_18, PPG, MPG, TOPG, RPG, PER, SPG, APG)
ggpairs(stats_salary_cor)
cor(stats_salary_cor)[,"salary17_18"]
```

## Data Visualization
### Interactive Plot
```{r}
names(stats_salary)[5] <- "Team"
plot_ly(data = stats_salary, x = ~salary17_18, y = ~PPG, color = ~Team,
        text = ~paste("Player: ", Player))
```

### Scatter plot with regression line
```{r}
stats_salary %>% 
  ggplot(aes(x = salary17_18, y = APG)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Regression Analysis
```{r}
stats_salary_regression <- 
  stats_salary %>% select(salary17_18, MPG:SPG)
lm(salary17_18~., data=stats_salary_regression)
stats_salary_regression %>% head()
```

1.Point per game increases salary by $686,815 per year
2.The more assists the more salary players get
3.Less turnovers is great


### Is the Player trusted by coach? How many turnovers does the player make?
```{r}
summary(stats_salary_regression$TOPG)
```

```{r}
avg.minutes <- mean(stats_salary_regression$MPG)
avg.turnover <- mean(stats_salary_regression$TOPG)
stats_salary_regression$Trusted <- as.factor(ifelse(stats_salary_regression$MPG >= avg.minutes, "Yes", "No"))
stats_salary_regression$Agressiveness <- as.factor(ifelse(stats_salary_regression$TOPG >= avg.turnover, "Yes", "No"))
head(stats_salary_regression)
```
### Scatter plot coloured by how many turnovers they make
```{r}
stats_salary_regression %>% 
  ggplot(aes(x = salary17_18, y = PPG, colour = Agressiveness)) + 
  geom_point() + 
  geom_smooth(method="lm")
```


```{r}
lm(formula = salary17_18 ~ Trusted * Agressiveness, data=stats_salary_regression)
```

if you are trusted by coach (which means, in this case, you get enough play time) and score, the salary increase by $325,079.

## Modeling & Conclusion
### Prediction Function
```{r}
salary_prediction <- function(m, point, minutes, turn_over){
  pre_new <- predict(m, data.frame(PPG = point, MPG = minutes, TOPG = turn_over))
  msg <- paste("PPG:", point, ",MPG:", minutes, ",TOPG:", turn_over, " ==> Expect Salary: $", round(pre_new), sep = "")
  print(msg)
}
```


### Analysis conclusion
```{r}
model <- lm(formula = salary17_18 ~ PPG + MPG + TOPG, data = stats_salary_regression)
salary_prediction(model, 23.7, 34.1	, 2.3)


# model <- lm(charges ~ age+bmi+children, data=insurance)
# pre_charges(model, 19, 27.9, 0)
```





