---
title: "NBA Players' Salary Prediction using liear regression model"
author: "Koki Ando"
date: "3/19/2018"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: cosmo
    toc: yes
    toc_depth: 3
    number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Hi, my name is Koki Ando. <br>
I've been a big fan of NBA for about 10 years. And recently, Ive been teaching myself R, and i really enjoy using it especcially when im dealing with data about basketball.<br>
Ive done some data analysis about basketball before, but this time, I'm going to dive into nba players' salary data. through this analysis, what i liked about was predicting something using linear regression model, which i just taught myself a few days ago. hope you guys also find this data analysis interesting.


# Preparation
## Required packages

```{r}
library(data.table)
library(corrplot)
library(GGally)
library(tidyverse)
library(PerformanceAnalytics)
library(plotly)
```

## Data Preparation
"salary.table" dataset was not provided at kaggle. So i scraped [this site](https://www.basketball-reference.com/contracts/players.html) and created by myself.
full code is on my [GitHub](https://github.com/koki25ando/NBA-Players-2017-18-dataset) page. let me know whatever u think.
I needed to create this dataset because i wanted to compare salary dataset of this season(17-18) to the nba players stats of the previous season(16-17). <br>
I wanted to investigate the how the stats effect next season's salary they get.

```{r}
salary.table <- 
  fread("https://s3-ap-southeast-2.amazonaws.com/koki25ando/NBA_season1718_salary.csv", 
        data.table = FALSE)
ss <- fread("https://s3-ap-southeast-2.amazonaws.com/playerinfomation/Seasons_Stats.csv",
            data.table = FALSE)
```
just in case u want to know how the datasets look like.
```{r}
str(salary.table)
str(ss)
```

### Data Cleaning
I wanted to use the stats data of 2016-17 only, so that i filtered. 
And unfortunattely stats per game was not in this "ss" dataset. so i mutated them all.
```{r}
stats17 <- 
  ss %>% filter(Year >= 2017) %>% 
  select(Year:G, MP, PER, FG:PTS) %>% 
  distinct(Player, .keep_all = TRUE) %>% 
  mutate(MPG = MP/G, PPG = PTS/G, APG = AST/G, 
         RPG = TRB/G, TOPG = TOV/G, BPG = BLK/G, 
        SPG = STL/G) 
```
### Merging Data
```{r}
stats_salary <- merge(stats17, salary.table, by.x = "Player", by.y = "Player")
names(stats_salary)[40] <- "salary17_18"
stats_salary <- stats_salary[-39]
```

# Correlation check
## Correlation check No.1
```{r}
corrplot(cor(stats_salary %>% 
               select(salary17_18, MPG:SPG, 
                      Age, PER, contains("%")), 
             use = "complete.obs"), 
         method = "circle",type = "upper")
```

## Correlation check No.2
```{r}
stats_salary_cor <- 
  stats_salary %>% 
  select(salary17_18, PPG, MPG, TOPG, RPG, PER, SPG, APG)
ggpairs(stats_salary_cor)
cor(stats_salary_cor)[,"salary17_18"]
```

Correlation strength is:
PPG > MPG > TOPG > RPG > PER > SPG > APG

the interesting part of this is that the number of turnover players make is linked to their salary, and they have a positive correlation.<br>
So, i interpreted this relationship like this: the more turnovers they makes means that they are more involved in ball movement, which means that players who make turnovers are, at some extend, important to their team. and i thought this could be expressed as "agressiveness". i know this interpretation could not be appropriate one. maybe next time, i should get data which include ho long players keep ball.

# Data Visualization
## Interactive Plot
```{r}
names(stats_salary)[5] <- "Team"
plot_ly(data = stats_salary, x = ~salary17_18, y = ~PPG, color = ~Team,
        text = ~paste("Player: ", Player))
```

If you hover on the points in this plot, you can check the player names. <br>
As you can tell the most paid player is Stephen Curry.
Last summer he got an extended contract with Warriors. 

## Scatter plot with regression line
```{r}
stats_salary %>% 
  ggplot(aes(x = salary17_18, y = PPG)) + 
  geom_point() + 
  geom_smooth(method = "lm") 
```

# Regression Analysis
```{r}
stats_salary_regression <- 
  stats_salary %>% select(salary17_18, MPG:SPG)
lm(salary17_18~., data=stats_salary_regression)
```

1. Point per game increases salary by $686,815 per year
2. The more assists they make the more salary they get


## Is the Player trusted by coach? How many turnovers do they make?

```{r}
avg.minutes <- mean(stats_salary_regression$MPG)
avg.turnover <- mean(stats_salary_regression$TOPG)
stats_salary_regression$Trusted <- as.factor(ifelse(stats_salary_regression$MPG >= avg.minutes, "Yes", "No"))
stats_salary_regression$Agressiveness <- as.factor(ifelse(stats_salary_regression$TOPG >= avg.turnover, "Yes", "No"))
head(stats_salary_regression)
```
## Scatter plot coloured by how many turnovers they make
```{r}
stats_salary_regression %>% 
  ggplot(aes(x = salary17_18, y = PPG, colour = Agressiveness)) + 
  geom_point() + 
  geom_smooth(method="lm")
```
As i mentioned before, players who get paid more tend to make more turnovers.

```{r}
lm(formula = salary17_18 ~ Trusted * Agressiveness, data=stats_salary_regression)
```

if you are trusted by coach (which means, in this case, you get play time more than average in the entire league) and score, the salary increase by $325,079.

# Modeling & Conclusion
## Prediction Function
```{r}
salary_prediction <- function(m, point, minutes, turn_over){
  pre_new <- predict(m, data.frame(PPG = point, MPG = minutes, TOPG = turn_over))
  msg <- paste("PPG:", point, ",MPG:", minutes, ",TOPG:", turn_over, " ==> Expected Salary: $", round(pre_new), sep = "")
  print(msg)
}
```

Let's take a example.<br>
In this season, as we all know Philadelphia 76ers are having a great season. This time i am going to fucus on JJ Redick.<br>
<br>
According to [Basketball Reference : J.J. Redick](https://www.basketball-reference.com/players/r/redicjj01.html), 
Redick's stat is "Points per game : 16.7, Minutes per game : 31.2, Turnovers per game : 1.5" so far in this season.<br>
And he doesn't have contract with 76ers for nect season.<br>
Let's predict.

## Analysis conclusion
```{r}
model <- lm(formula = salary17_18 ~ PPG + MPG + TOPG, data = stats_salary_regression)
salary_prediction(model, 16.7, 31.2, 1.5)
```

According to my prediction model, he will get $13,959,120 next season.

